Large-scale harmonic patterns clustering

### 2010-03-21 Paper/Paper.tex
improved

@@

\subsection{USPOP}
We had acces to a low quality (32kbps) version of the songs in the uspop 2002 
data set \cite{uspop2002}.
This well-known data show a great diversity of pop songs, but is not particular
otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
not contain any bar.

@@

\subsection{Artist Set}
This set ecompass $1402$ songs $20$ artists, mostly rock and pop of different 
subgenres. Some results were available regarding the artist recognition
task. Few details given to preserve anonymity.

@@

\subsection{Pattern Creation}
Here we should explain key invariance, resizing, etc, and not in the
experiment section.

We try to normalize the patterns in respect to the key. We roll the matrix
so the row with the more energy is always the lowest one. This can easily be
seen in Figure \ref{fig:code}. We considered the same technique for rolling
along the other axis and be invariant regarding the downbeat. However, it
seems less robust to noise. Note that we roll each pattern independentely.
Also, this implies that the pitch envelope is important rather than the actual 
pitches.

@@

These properties are important to verify and measure, but they are not
surprising for a clustering algorithm. We move on to more interestng facts.

@@

\subsection{Artist Recognition}
Let us be clear, the clustering method proposed in this work is not
intended to perform well at artist recognition. An obvious fact is that
we lose all timbral information which would be usefull. However, we can
imagine that some artist use more often some patterns than others.
Best accuracy reported\footnote{few details given to preserve anonymity} 
is around $54.5\%$. Our result is $23.4\%$. Random
is around $5\%$. The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
Details of the experiment follow.

>>>>

We approach the clustering problem keeping in mind the scaling
requirement. We use a representation similar to the \textit{Shingles}.
We train on $43$K songs and test on more than $7$K ones.

@@

\subsection{USPOP}
We had access to a low quality (32kbps) version of the songs in the uspop 2002 
data set \cite{uspop2002}.
This well-known data show a great diversity of pop songs, but is not particular
otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
not contain any bar.

@@

\subsection{Artist Set}
This set encompass $1402$ songs $20$ artists, mostly rock and pop of different 
subgenres. Some results were available regarding the artist recognition
task. Few details given to preserve anonymity.

@@

\subsection{Pattern Creation}
Patterns are defined as a number of beats or bars. In the case of beats,
we simply cut the set of chromas per beat every $N$ beats. In the case
of bars, we gather all beats corresponding to the $N$ bars requested.
The number of chroma vectors we get can vary. We resize the features in the
time axis to a fixed length. Resizing can be argued over, but it should
preserve the main bar characteristics (such as a strong beat at the beginning
or a note change in the middle).

We try to normalize the patterns in respect to the key. We roll the matrix
so the row with the more energy is always the lowest one. This can easily be
seen in Figure \ref{fig:code}. We considered the same technique for rolling
along the other axis and be invariant regarding the downbeat. However, it
seems less robust to noise. Note that we roll each pattern independently.
Also, this implies that the pitch envelope is important rather than the actual 
pitches.

@@

These properties are important to verify and measure, but they are not
surprising for a clustering algorithm. We move on to more interesting facts.

@@

\subsection{Artist Recognition}
Let us be clear, the clustering method proposed in this work is not
intended to perform well at artist recognition. An obvious fact is that
we lose all timbral information which would be useful. However, we can
imagine that some artist use more often some patterns than others.
Best accuracy reported\footnote{few details given to preserve anonymity} 
is around $54.5\%$. Our result is $23.4\%$. Random
is around $5\%$. The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
Details of the experiment follow.
